---
title: "hw03"
author: "Ting An Lai"
date: '2019-10-06'
output:
#  html_document:
#    df_print: paged
#    toc: yes
  html_notebook:
    toc: yes
---

## 1. Monkey Song
### a) Write the loop: 
1) iterated from 10 to the last verse
2) print 10 to 8 monkeys to console
```{r}
# Create a df storing the verses
song.df <- vector("character", 11)

# for loop that 1) iterates from 10 to the last verse 2) print 10 to 8 monkeys to console
for (i in 10:0) {
  
  #Iterate all monkey verses into the song dataframe
  if (i > 1){
    song.df[11-i] <- paste (i, 'little monkeys jumping on the bed, One fell down and
                            bumped his head,Mama called the doctor and the doctor
                            said,"No more monkeys jumping on the bed!"')
    } else if (i == 1) {
      song.df [11-i] <- c('1 little monkey jumping on the bed, He fell down and bumped
                          his head, Mama called the doctor and the doctor said,"No
                          more monkeys jumping on the bed!"')
      } else {
        song.df [11-i] <- c('No little monkeys jumping on the bed,None fell down an
                            bumped their heads,Papa called the doctor and the doctor
                            said,"Put those monkeys right to bed!"')
        }
  
  #Only print out 10 to 8 monkeys verse
  if (i >= 8) {
    cat (i, 'little monkeys jumping on the bed,\nOne fell down and bumped his head,\nMama called the doctor and the doctor said,\n"No more monkeys jumping on the bed!"\n\n')
    } 
}
```


### b) c) Convert loop to a function

#### Function
```{r}
animal.song <- function (animal = "monkey", number = 10){
  animal.plur = paste (animal, "s", sep = "")
  
  # Warning message for numeric entry
  if (is.numeric(number) == FALSE) {
    print ("Please enter a numeric number for the amount of animals")
  } else if (number <3) {
    print ("Please enter a number larger than 3")
  } else {
 
  # Create empty vector for later printing
  song.df <- vector ("character", number)
  
  # Looping for different iterations
  for (i in number:0) {
    if (i > 1){
      song.df[number-i+1] <-paste (i, 'little', animal.plur, 'jumping on the bed,\nOne fell down and bumped his head,\nMama called the doctor and the doctor said,\n"No more', animal.plur, 'jumping on the bed!"\n\n')
      } else if (i == 1) {
        song.df[number-i+1] <-  paste('1 little', animal, 'jumping on the bed,\nHe fell down and bumped his head,\nMama called the doctor and the doctor said,\n"No more', animal.plur,'jumping on the bed!"\n\n')
        } else {
          song.df[number-i+1] <- paste('No little',animal.plur, 'jumping on the bed,\nNone fell down and bumped their heads,\nPapa called the doctor and the doctor said,\n"Put those',animal.plur,'right to bed!"\n\n\n')
        }
    }
  cat (song.df)
  }
  
}
```

#### Default animal, monkeys; default number, 10
```{r}
animal.song ()
```

#### 3 Piglets verse
```{r}
animal.song("Piglet", 3)
```

### Warning message testing
```{r}
animal.song ("Cats", TRUE)
```

### d) BONUS (optional): Convert loop to a fully generalized function
```{r}
animal.song.gen <- function (animal = "monkey", number = 10, activity = "jumping", parent.prob = 0.5){
  # Modify animal plurality
  animal.plur = paste (animal, "s", sep = "")
  # Address mama papa probability 
  parent.vector = rbinom (number+1, 1, parent.prob)
  parent.vector <- gsub (1, "Mama", parent.vector)
  parent.vector <- gsub (0, "Papa", parent.vector)

  # Address incorrect input warning signs
  
  if (is.numeric(number) == FALSE) {
    print ("Please enter a numeric number for the amount of animals")
  } else {
  # Create empty vector for later printing
  song.df <- vector ("character", number)
  
  # Looping for different iterations
  for (i in number:0) {
    if (i > 1){
      song.df[number-i+1] <-paste (i, 'little', animal.plur, activity,  'on the bed,\nOne fell down and bumped his head,\n', parent.vector[number-i+1], 'called the doctor and the doctor said,\n"No more', animal.plur, activity, 'on the bed!"\n\n')
      } else if (i == 1) {
        song.df[number-i+1] <-  paste('1 little', animal, activity, 'on the bed,\nHe fell down and bumped his head,\n', parent.vector[number-i+1], 'called the doctor and the doctor said,\n"No more', animal.plur,activity, 'on the bed!"\n\n')
        } else {
          song.df[number-i+1] <- paste('No little',animal.plur, activity, 'on the bed,\nNone fell down and bumped their heads,\n', parent.vector[number-i+1], 'called the doctor and the doctor said,\n"Put those',animal.plur,'right to bed!"\n\n\n')
        }
    }
  cat (song.df)
  }
  
}

```

```{r}
animal.song.gen("kangaroo", 3, "dancing", 0.5)
```


## 2. Draw a plot using a loop

Let's draw a plot with a loop. You can inialize a new plot with plot(). Lines can be added to a plot with lines() command. For example, lines(x = c(1,1), y = c(1,7)) will draw a line between the coordinates (1,1) and (1,7).

### a) b) c) Plot horizontal,  vertical lines, and points
```{r}
plot(1, type="n", xlab="", ylab="", xlim = c(0, 10), ylim = c(0, 10))
for (i in seq (0,10)) {
  lines(x = c(0, 10), y = c(i,i))
  lines (x = c(i, i), y = c(0, 10))
}
for (i in seq (0, 10)) {
  points (x = i, y = i)
  points (x = i, y = 10-i)
}

```

### d) Convert to function
```{r}
plot.draw <- function (size = 10, horizontal_line = TRUE, vertical_line = TRUE, points_on_diagonal = TRUE ) {
  plot(1, type="n", xlab="", ylab="", xlim = c(0, size), ylim = c(0, size))
  if (horizontal_line){
    for (i in seq (0,size)) {
    lines(x = c(0, size), y = c(i,i))
    }
  }  
  if (vertical_line) {
    for (i in seq (0,size)) {
    lines(x = c(i, i), y = c(0,size))
    }
  }  
  if (points_on_diagonal) {
    for (i in seq (0, size)) {
  points (x = i, y = i)
  points (x = i, y = size-i)
}
}
}


#plot.draw ()
#plot.draw (5, TRUE, TRUE, FALSE)
```



